---
title: "Summary Statistics and Plots"
output:
  html_document:
    df_print: paged
---

In this notebook we will load in the Palmer penguins dataset, and your task is to show how R can be used for summary statistics and plots, so that we can understand the key features of the dataset, and the differences between penguin species.

In order to install the package containing the data file you will need to run `install.packages("palmerpenguins")` in the R console window. (We do not include the install command in the Notebook because we only run it once, not everytime we run the Notebook).

```{r}
library("tidyverse")
library("palmerpenguins")
```

In groups work to add example code (and any appropriate comments on when a statistic/plot is a good/bad choice) for each of the headings below.

## Summary Statistics

```{r}
summary(penguins)
```

### Measures of central tendancy

```{r}
head(penguins)
```

**mean**

```{r}
penguins %>%
  select(bill_length_mm:body_mass_g) %>%
  drop_na()%>%
  summarise_all(mean)
```

**median**

```{r}
penguins %>%
  select(bill_length_mm:body_mass_g) %>%
  drop_na()%>%
  summarise_all(median)
```

### Measures of dispersion

**standard deviation**

```{r}
penguins %>%
  select(bill_length_mm:body_mass_g) %>%
  drop_na()%>%
  summarise_all(sd)
```

**variance**

```{r}
penguins %>%
  select(bill_length_mm:body_mass_g) %>%
  drop_na()%>%
  summarise_all(var)
```

**range**

```{r}
penguins %>%
  select(bill_length_mm:body_mass_g) %>%
  drop_na()%>%
  summarise_all(range)
```

**inter-quartile range**

```{r}
penguins %>%
  select(bill_length_mm:body_mass_g) %>%
  drop_na()%>%
  summarise_all(IQR)
```

## Visualising Data

### Visualising categorical data

**barcharts**

```{r}
ggplot(penguins, aes(x=species)) +
  geom_bar() +
  labs(title = "Count of Penguins by Species")
```

**frequency table**

```{r}
table(penguins$species, penguins$island, penguins$sex)
```

### Visualising distributions

**histograms**

```{r}
ggplot() +
  geom_histogram(data=filter(penguins),
                 mapping = aes(x=bill_length_mm),
                 binwidth = 5)
```

**density plots**

```{r}
ggplot(data = penguins %>% drop_na(sex, bill_depth_mm)) +
  geom_density(mapping=aes(x=bill_depth_mm, color=sex))
```

## Comparisons

### two continuous variables

**scatter plot**

```{r}
ggplot(data=penguins %>% drop_na(sex, bill_length_mm, bill_depth_mm), mapping = aes(x=bill_depth_mm, y = bill_length_mm, color=sex)) +
  geom_point(size=0.1, alpha=0.8) 

ggplot(data=penguins, mapping = aes(x=bill_depth_mm, y = bill_length_mm, color=species)) +
  geom_point(size=0.1, alpha=0.8) 
```

**scatter plot with smooth trendline**

```{r}
ggplot(data=penguins, mapping = aes(x=bill_depth_mm, y = bill_length_mm, color=species)) +
  geom_point(size=0.1, alpha=0.8) +
  geom_smooth()
```

**scatter plot with straight line fit**

```{r}
ggplot(data=penguins, mapping = aes(x=bill_depth_mm, y = bill_length_mm, color=species)) +
  geom_point(size=0.1, alpha=0.8) +
  geom_smooth(method="lm")

```

**2D density plots**

```{r}
ggplot(data=penguins,mapping=aes(x=bill_depth_mm, y=bill_length_mm)) +
  geom_density2d()
```

**pairs plot**

```{r}
pairs(penguins %>% select(bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g), col = penguins$species)

```

**correlation**
```{r}
penguins %>%
  ggcorr(label=TRUE)
```


```{r}
cor(penguins %>% select(bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g) %>% drop_na())

```

### continuous vs categorical variables

**histograms**
```{r}
penguins %>%
  drop_na(bill_length_mm) %>%
  ggplot(aes(bill_length_mm)) +
  geom_histogram() +
  facet_grid(cols=vars(species))
```

```{r}
ggplot(data = penguins, mapping = aes(x=bill_depth_mm, y=species)) +
  geom_histogram()
```

**density plots**
```{r}
penguins %>%
  drop_na(bill_length_mm) %>%
  ggplot(aes( x=bill_length_mm, fill=species) )+
  geom_density(alpha=0.2, position="identity") 
```

**violin plots**
```{r}
penguins %>%
  drop_na(bill_length_mm) %>%
  ggplot(aes(y=bill_length_mm, x=species)) +
  geom_violin()
```

**box plots**

```{r}
penguins %>%
  drop_na(bill_length_mm) %>%
  ggplot(aes(y=bill_length_mm,x=species)) +
  geom_boxplot()

```

### two categorical variables

**frequency table**

```{r}
penguins %>%
  count(species,island) %>%
  pivot_wider(names_from=species,values_from=n,values_fill=0)
```

