|                       |
|-----------------------|
| title: "Day 3"        |
| output: html_notebook |

## Introduction to R Notebook

R Notebooks provide a powerful and flexible way to interactively develop and share your R code. They allow you to combine code, output, and narrative text in a single document, making them ideal for data analysis, exploration, and reporting.

. Getting Started with R Notebooks

1- Creating an R Notebook

-   Open RStudio.

-   Go to `File > New File > R Notebook`.

2- Structure of an R Notebook

-   An R Notebook consists of text chunks and code chunks.

-   Text chunks are written in Markdown.

-   Code chunks are written in R and are enclosed by triple back ticks with `{r}` at the beginning.

-   You can run the code chunks individually and see the output immediately below the chunk

## Simple code block in R Notebook

```{r}
#This is a code chunk 
print("Hello")
```

-   Follow this link for more information on [R Markdown](http://rmarkdown.rstudio.com) .

## Using R Notebook and Tidyverse function for `hsb2` dataset

-   We will address some questions from the Day 3 document by utilizing tidyverse functions and documenting our code within an R Notebook.

```{r, warning=FALSE,error=FALSE}

library(tidyverse)
```

```{r}
#Import Highschool data  and call it "highschool_raw"
highschool_raw = read_csv("/Users/pp16/Desktop/NTU/UCL exchange/week 1 day 3/hsb2.csv")
```

### changing the levels enteries to labels using `mutate`

```{r}

#use Mutate command(change the levelss to factor variables)  to creat a dataframe named "highschool" from highschool_raw 
library(dplyr)
highschool = mutate(highschool_raw,
                    gender = factor(female, levels=c(0,1), labels=c("M","F")),
                    race = factor(race, levels=c(1,2,3,4), labels = c("hispanic", "asian", "african-american", "white")),
                    ses = factor(ses,levels = c(1,2,3), labels = c("low", "middle", "high")),
                    schtyp = factor(schtyp, levels=c(1,2), labels=c("public","private")),
                    prog = factor(prog, levels=c(1,2), labels=c("general","academic" )),
)


#remove female column, use `select` command
highschool_modified = select(highschool, race, ses, schtyp, prog, read, write, math, science, socst)
highschool_modified = select(highschool, -female)

# check highschool data 
head(highschool)
head(highschool_modified)
```

### renaming columns with `rename`

```{r}

#rename columns as follows:
#socioeconomic_group = ses,
#racial_group = race,
#school_type = schtyp,

highschool_rename = rename(highschool_modified, 
                           socioeconomic_group = ses,
                           racial_group = race,
                           school_type = schtyp,)

head(highschool_rename)
```

### Examples of tidyverse code

-   Answer the following : **Average scores for girls from public school**

```{r}
scores = filter(highschool, schtyp=="public", gender=="F")
avgscore = select(scores, read, write, math, science, socst)
meanscore = summarise(avgscore,
                      avgread= mean(read),
                      avgwrite=mean(write),
                      avgmath = mean(math),
                      avgscience=mean(science),
                      avgsocst=mean(socst),)
print(meanscore)
```

**Average scores for African-american students**

```{r}
a = filter(highschool, race=="african-american")
newa = select(scores, read, write, math, science, socst)
meana = summarise(avgscore,
                      avgread= mean(read),
                      avgwrite=mean(write),
                      avgmath = mean(math),
                      avgscience=mean(science),
                      avgsocst=mean(socst),)
print(meana)

```

**Data frame containing only student scores**

```{r}
student = select(scores, id, read:socst)
student
```

### Pipes

Pipes `%>%` allow you to chain commands together, making your code more readable and efficient. The `tidyverse` package provides the pipe operator, and it is also included in the `dplyr` package.

## Pipe operators : `hsb2` data

-   Now that we are familiar with the hsb2 data, let's use the pipe operator to make our code more efficient and avoid creating multiple data subsets.

-   The following code demonstrates how to use the pipe operator to mutate, remove a column, and rename variables in the dataset, applying all these changes efficiently in a single pipeline:

-   In this code:

-   `mutate(new_column = some_transformation(existing_column))`: Adds a new column by transforming an existing one.

-   `select(-unwanted_column)`: Removes the specified column.

-   `rename(new_name1 = old_name1, new_name2 = old_name2)`: Renames specified columns.

-   Using the pipe operator (%\>%), all these operations are performed sequentially and the resulting dataset is assigned back to hsb2.

```{r}
highschool <- read_csv("/Users/pp16/Desktop/NTU/UCL exchange/week 1 day 3/hsb2.csv") %>% 
  mutate(race = factor(race, levels = c(1,2,3,4),
                             labels = c("hispanic", 
                                        "asian", 
                                        "african-american", 
                                        "white")),
         ses = factor(ses, levels = c(1,2,3),
                           labels = c("low", 
                                      "middle", 
                                      "high")),
         schtyp = factor(schtyp, levels = c(1,2),
                           labels = c("public", 
                                      "private")),
         gender = factor(female, levels = c(0,1),
                           labels = c("male", 
                                      "female"))) %>% 
  select(-female) %>% 
  rename( socioeconomic_group = ses, 
          racial_group = race, 
          school_type = schtyp)

highschool
```

-   write a code to provides a score_summary for females who are from public schools using the pipe operator to filter, summarize, and format the data:

## Your turn!

-   write a code to provides a score_summary for females who are from public schools using the pipe operator to filter, summarize, and format the data:

-   You may use the following commands:

-   Filtering [[[[[[Data:\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\`filter(gender](Data:\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\%60filter(gender){.uri}]([Data:\\\\\\\\\\\\\\\\\\\\\\\\\\\\\\%60filter(gender)](Data:\\\\\\\\\\\\\\\%60filter(gender)){.uri}{.uri}]([Data:\\\\\\\\\\\\\\%60filter(gender)](Data:\\\%60filter(gender)){.uri}{.uri}]([Data:\\\\\\%60filter(gender)](Data:%60filter(gender)){.uri}{.uri}]([Data:\\%60filter(gender)](Data:%60filter(gender)){.uri}{.uri}]([Data:%60filter(gender)](Data:%60filter(gender)){.uri}{.uri} == "female", school_type == "public")\`: Selects rows where the gender column is "female" and the school_type column is "public".

-   `select(read:socst)`: Selects columns from read to socst. These columns typically represent different subjects such as reading, math, science, and social studies.

-   `summarise_all(mean)`: Computes the mean (average) for each selected column. This operation calculates the average score for each subject across all female students in public schools.

-   After executing this code, `score_summary` will contain a single row of data with the average scores for reading, math, science, and social studies among female students attending public high schools.

```{r}

score_summary = highschool %>%
  filter(gender=="female", school_type=="public") %>%
  select(read:socst)%>%
  summarise_all(mean)

score_summary

```

### Missing data

```{r}
highschool_missing <- read_csv("/Users/pp16/Desktop/NTU/UCL exchange/week 1 day 4/hsb2_missing_data.csv") %>% 
  mutate(race = factor(race, levels = c(1,2,3,4),
                             labels = c("hispanic", 
                                        "asian", 
                                        "african-american", 
                                        "white")),
         ses = factor(ses, levels = c(1,2,3),
                           labels = c("low", 
                                      "middle", 
                                      "high")),
         schtyp = factor(schtyp, levels = c(1,2),
                           labels = c("public", 
                                      "private")),
         gender = factor(female, levels = c(0,1),
                           labels = c("male", 
                                      "female"))) %>% 
  select(-female) %>% 
  rename( socioeconomic_group = ses, 
          racial_group = race, 
          school_type = schtyp)

summary(highschool_missing)
```

```{r}
score_missing = highschool_missing %>%
  drop_na() %>%
  filter(gender=="female", school_type=="public") %>%
  select(read:socst)%>%
  summarise_all(mean)
  
score_missing
```

```{r}
mean(highschool_missing$math, na.rm=T)
```

# See number of complete cases

```{r}
sum(complete.cases(highschool_missing))
```
